---
title: "Interactive Satellite Map"
author: "Caitlyn Gramajo"
date: "2024-04-02"
output: 
  html_document:
    code_folding: hide
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Idea

One of my favorite things to do it make cool pictures so the goal of my project was simple to make a cool picture. I asked one of my cousins who is obsessed with space if he knew of good space data I could play with. He directed me to a place with satellite data and the idea came to me. I was going to make a 3 dimensional plot of the satellites around earth that could be moved with a curser.

## Data source

The data I am using is from https://www.ucsusa.org/resources/satellite-database

## Plot

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(rgl) 

dat<-read_xlsx("UCS-Satellite-Database 5-1-2023.xlsx")
dat1<-clean_names(dat)
#colnames(dat1)

## Dropping everything past comments because I don't know what it is and it is mostly NA ####

selected_columns <- colnames(select(dat1, c(1:26)))

dat2<-
  dat1 %>% 
  select(all_of(selected_columns))

#colnames(dat2)

## I only want orbit stuff ####
selected_columns2 <- colnames(select(dat2, c("current_official_name_of_satellite","country_org_of_un_registry","country_of_operator_owner","operator_owner" ,"users","purpose","class_of_orbit","type_of_orbit","longitude_of_geo_degrees","perigee_km","apogee_km","eccentricity","inclination_degrees","period_minutes")))

dat3<-
  dat2 %>% 
  select(all_of(selected_columns2))

## Check for NAs in all columns ####
#sapply(dat3, function(col) any(is.na(col)))

# Drop rows with NAs
dat4 <- na.omit(dat3)

## Making perigee + apogee numeric ####

dat5 <- dat4 %>%
  mutate(perigee = as.numeric(perigee_km),  # Convert 'perigee' to numeric
         apogee = as.numeric(apogee_km)) 

## Function for satellite posititon #####
calculate_satellite_positions <- function(longitude_of_geo_degrees, perigee_km, apogee_km, eccentricity, inclination_degrees, period_minutes) {
  # Constants
  R_earth <- 6371  # Earth radius in km
  
  # Initialize lists to store coordinates
  x <- y <- z <- NULL
  
  # Convert inputs to numeric
  longitude_of_geo_degrees <- as.numeric(longitude_of_geo_degrees)
  perigee_km <- as.numeric(perigee_km)
  apogee_km <- as.numeric(apogee_km)
  eccentricity <- as.numeric(eccentricity)
  inclination_degrees <- as.numeric(inclination_degrees)
  period_minutes <- as.numeric(period_minutes)
  
  # Check for non-numeric values
  if (any(is.na(longitude_of_geo_degrees)) || any(is.na(perigee_km)) || any(is.na(apogee_km)) ||
      any(is.na(eccentricity)) || any(is.na(inclination_degrees)) || any(is.na(period_minutes))) {
    stop("Input arguments must be numeric")
  }
  
  for (i in seq_along(longitude_of_geo_degrees)) {
    # Orbital parameters for current satellite
    lon <- longitude_of_geo_degrees[i]
    perigee <- perigee_km[i]
    apogee <- apogee_km[i]
    ecc <- eccentricity[i]
    incl <- inclination_degrees[i]
    period <- period_minutes[i]
    
    # Semi-major axis (average of perigee and apogee)
    semi_major_axis <- (perigee + apogee) / 2
    
    # Orbital period in seconds
    period_seconds <- period * 60
    
    # True anomaly
    M <- 2 * pi * (period_seconds / 86400)  # Mean anomaly
    E <- M  # Initial guess for eccentric anomaly
    iter <- 0
    while (iter < 100) {  # Limit iterations to prevent infinite loop
      E_next <- E - (E - ecc * sin(E) - M) / (1 - ecc * cos(E))
      if (abs(E_next - E) < 1e-8) break  # Convergence criteria
      E <- E_next
      iter <- iter + 1
    }
    
    # Distance from center to focus
    r <- semi_major_axis * (1 - ecc * cos(E))
    
    # True anomaly
    theta <- atan2(sqrt(1 - ecc^2) * sin(E), cos(E) - ecc)
    
    # Convert orbital elements to 3D coordinates
    x <- c(x, r * (cos(lon) * cos(incl) * cos(theta) - sin(lon) * sin(theta)))
    y <- c(y, r * (sin(lon) * cos(incl) * cos(theta) + cos(lon) * sin(theta)))
    z <- c(z, r * sin(incl) * sin(theta))
  }
  
  return(data.frame(x = x, y = y, z = z))
}



## Calculate satellite positions with function ####
satellite_positions <- calculate_satellite_positions(
  dat5$longitude_of_geo_degrees,
  dat5$perigee_km,
  dat5$apogee_km,
  dat5$eccentricity,
  dat5$inclination_degrees,
  dat5$period_minutes
)

## Plotting ####

# Create a 3D plot window
open3d()

# Create a 3D sphere representing the Earth
rgl::spheres3d(0, 0, 0, radius = 6371, color = "blue")

# Plot satellite positions on the sphere
rgl::points3d(satellite_positions$x, satellite_positions$y, satellite_positions$z, col = "darkorchid3", size = 5)

# Add axis labels
rgl::text3d(100000, 0, 0, "X", adj = c(-0.5, 0.5), col = "black")
rgl::text3d(0, 100000, 0, "Y", adj = c(-0.5, 0.5), col = "black")
rgl::text3d(0, 0, 100000, "Z", adj = c(-0.5, 0.5), col = "black")

# Add lines from the origin to each axis
rgl::lines3d(c(-100000, 100000), c(0, 0), c(0, 0), col = "black")  # X-axis
rgl::lines3d(c(0, 0), c(-100000, 100000), c(0, 0), col = "black")  # Y-axis
rgl::lines3d(c(0, 0), c(0, 0), c(-100000, 100000), col = "black")  # Z-axis

# Add axes with scale lines and labels
axes3d(edges = c("x", "y", "z"), tick = TRUE, labels = TRUE,
       xlim = c(-100000, 100000), ylim = c(-100000, 100000), zlim = c(-100000, 100000))

# Save the 3D plot as an object
plot3d_object <- rgl::rglwidget()

plot3d_object

```

